---
title: "Paired data, replay and honest denominators"
output: html_document
---

A simulation comparison is easier to inspect when every adapter fits the same generated data and every attempt remains visible. `groundtruth` starts with Gaussian linear data and two coefficient targets. This local candidate adds an explicit logistic family.

```{r setup, include=FALSE}
library(groundtruth)
identity_inputs <- c(
  source_sha = Sys.getenv("GROUNDTRUTH_DOCS_EXPECTED_SOURCE_SHA"),
  package_path = Sys.getenv("GROUNDTRUTH_DOCS_EXPECTED_PACKAGE_PATH"),
  package_version = Sys.getenv("GROUNDTRUTH_DOCS_EXPECTED_PACKAGE_VERSION"),
  receipt_path = Sys.getenv("GROUNDTRUTH_DOCS_IDENTITY_RECEIPT")
)
if (any(nzchar(identity_inputs))) {
  stopifnot(all(nzchar(identity_inputs)))
  expected_hash <- identity_inputs[["source_sha"]]
  expected_path <- identity_inputs[["package_path"]]
  expected_version <- identity_inputs[["package_version"]]
  receipt_path <- identity_inputs[["receipt_path"]]
  namespace <- asNamespace("groundtruth")
  namespace_path <- normalizePath(getNamespaceInfo(namespace, "path"), mustWork = TRUE)
  stopifnot(identical(namespace_path, normalizePath(expected_path, mustWork = TRUE)))
  installed_identity <- dget(file.path(namespace_path, "source-identity.R"))
  stopifnot(identical(installed_identity$source_sha256, expected_hash),
            identical(installed_identity$package_version, expected_version))
  writeLines(capture.output(dput(list(
    namespace_path = namespace_path,
    source_sha256 = installed_identity$source_sha256,
    package_version = installed_identity$package_version,
    checks_passed = TRUE
  ))), receipt_path)
  message("DOCS_ARTICLE_SOURCE_IDENTITY_OK")
}
```

## Generate once, fit separately

```{r study}
scenario <- gt_scenario("small", n = 30, alpha = 1, beta = 2, sigma = 1)
study <- gt_runstudy(scenario, reps = 4, seed = "20261004")
head(study$data)
study$ledger
```

The Gaussian generator draws `x` and error independently. It constructs `y` from the stated truth without calling an estimator. Each adapter receives a fresh data frame containing only `x` and `y`. The runner stores truth separately for scoring.

## Fit a logistic scenario

For Bernoulli outcomes, the model is

$$
Y_i \sim \operatorname{Bernoulli}(p_i), \qquad
\operatorname{logit}(p_i)=\alpha+\beta x_i.
$$

The coefficients are conditional log odds. Select the family and native R adapter explicitly:

```{r logistic-study}
logistic <- gt_scenario("logistic-small", n = 80, alpha = -0.3, beta = 0.6,
                        family = "logistic")
logistic_study <- gt_runstudy(logistic, adapters = gt_glm_adapter(),
                              reps = 4, seed = "20261006")
head(logistic_study$ledger)
```

The native adapter fits `stats::glm()` with a binomial logit link. It requires native GLM convergence, a maximum absolute standardized score no greater than `1e-9`, and a maximum absolute standardized Newton correction to the linear predictor no greater than `1e-9`. Complete or quasi separation means the declared model has no unique finite MLE. A data set with a finite MLE can still fail numerical convergence, so the two conditions are reported separately. Gaussian fits retain Student t intervals; logistic fits use normal Wald intervals at 90% and 95%.

A custom logistic fitter must opt in with `gt_adapter(..., family = "logistic")`. Family compatibility is checked before calling the fitter. A mismatch is recorded as a failed attempt, and the fitter is not invoked. The native adapter is requested with `gt_glm_adapter()`.

## Keep points when intervals are absent

A custom point-only adapter can illustrate interval accounting:

```{r custom}
point_only <- gt_adapter("point-only", function(data) {
  fitted <- stats::lm(y ~ x, data = data, na.action = stats::na.fail,
                      singular.ok = FALSE)
  coefficient <- stats::coef(fitted)
  gt_result(c(alpha = unname(coefficient["(Intercept)"]),
              beta = unname(coefficient["x"])))
})
paired <- gt_runstudy(scenario,
  adapters = list(gt_adapter(), point_only), reps = 4, seed = "20261004")
summary <- gt_summarize(paired)
summary[, c("adapter", "target", "level", "attempted", "successful",
            "usable_intervals", "coverage", "covered_per_attempt")]
```

The point-only adapter still contributes to bias and RMSE. Its intervals are unusable, so conditional coverage is `NA` and covered per attempt is zero. A failed fit rejects both coefficient targets. Failure counts remain in the attempt denominator.

An interval result uses these exact column names:

```{r intervals}
gt_result(c(alpha = 1, beta = 2),
          intervals = data.frame(target = c("alpha", "beta"),
                                 level = c(0.95, 0.95),
                                 lower = c(0.5, 1.5),
                                 upper = c(1.5, 2.5)))
```

Missing intervals at the other level remain unusable. A duplicate target-level row does not select one row arbitrarily. Nonfinite or reversed bounds also exclude that interval while retaining valid points.

## Distinguish two kinds of interval

The default Gaussian fitted-coefficient intervals use Student t quantiles with `n - 2` degrees of freedom and residual variance `SSE / (n - 2)`. Logistic intervals use normal Wald quantiles and final observed information. The coverage summary uses a Wilson 95% Monte Carlo interval for the measured coverage fraction. The first interval describes coefficient uncertainty in one data set; the second describes uncertainty from a finite number of repetitions.

With four repetitions, the Monte Carlo uncertainty is large. This article demonstrates accounting and replay, not a calibration result.

## Retain replay information

```{r replay}
replayed <- gt_replay(paired)
stopifnot(identical(paired$data, replayed$data))
stopifnot(identical(paired$targets, replayed$targets))
head(paired$seed_map)
```

Replay uses the recorded RNG settings and full allocated seed map. Stateless deterministic adapters replay exactly in the same environment. For adapters that read external files or mutate captured state, preserve those inputs and reset that state yourself.

```{r export}
path <- tempfile("groundtruth-export-")
gt_export(paired, path)
manifest <- dget(file.path(path, "manifest.R"))
list.files(path)
```

A manifest is an audit record. `dget()` should be used only on a trusted manifest, because it evaluates R expressions. The exported manifest contains data rather than adapter code. It records a SHA256 for the sole R implementation file that produced the study and a separate exporter hash. This source identity does not include custom adapter code or the whole documentation. Read CSV columns according to their schemas, preserving seed strings as character.

R and Julia use distinct streams, so compare them on identical frozen CSV inputs. The
fixed-effect fixture results do not establish calibration or package-wide parity. R 4.6.0
checked version 0.0.0.9000: 346 assertions passed with `Status: OK`; the pkgdown 2.2.0
preview rendered. The candidate remains local and uncommitted. Publication approval has
not been given.
